SHARP - II: Revealing a bias in observational measurements of dark matter substructure with gravitational lens flux ratios

### 2015-11-04 papers/B1555_EdgeOn/B1555_letter.tex
text editing to modelling and discussion

The best-fit model parameters and 68\% CL errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:model}. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux-ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux-ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in fluxes may due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the exponential disc component makes up about 10 per cent of the total mass within the Einstein radius (1.7~kpc for a lens redshift of $z_l = 1$), which is not unreasonable. We note that there is a significant 0.036~arcsec offset between the SIE and exponential disc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models that use the full extent of the gravitational arc and the compact lensed images from future high resolution radio imaging will be able to confirm this.

@@

Our results for B1555+375 clearly demonstrate that a lack of knowledge about the lens galaxy mass distribution, combined with the limited information provided by the position of compact lensed images, can lead to improper lens models and a misinterpretation of the origin of flux-ratio anomalies. We've seen 5 out of 25 SHARP lenses revealing a disc in AO imaging, which these lenses are source selected thus less possible to be biased (Fassnacht et al., in prep.). While upcoming large-scale surveys with the Dark Energy Survey, Kilo-Degree Sky-VIKING, Square Kilometre Array, {\it Euclid}, and Large Synoptic Survey Telescope, are expected to deliver new large samples of gravitationally lensed quasars, in the near future the discovery of new systems using narrow-line quasar emission \citep{N14} will provide a new and invaluable statistical perspective on the current constraints based on flux-ratio anomalies. We stress, therefore, that one should be very careful when interpreting the nature of flux-ratio anomalies, as prior simple assumptions on the lens mass distribution may a significant source of bias.

Our analysis shows that in order to derive a precise quantification of the substructure population and other properties from strongly lensed flux-ratio anomalies, it is crucial to have a careful estimation of non-substructure effects, such as galaxy discs and nearby massive satellites. In fact, even in the presence of substructures, the observed flux-ratio anomalies could be still affected by the complex structure of the lens galaxy. Finally, additional near-infrared AO imaging from SHARP shows that B1555+375 is not the only gravitational lens having a significant disc component detection in the lensing galaxy.  These gravitational lenses will be investigated in a future work, but the images alone strongly suggest that other systems may have similar links between flux-ratio anomalies and disc components.

>>>>

The best-fit model parameters and 68\% CL errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:model}. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux-ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux-ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in fluxes may due to the real lensed source being slightly extended rather than the point source that we used in our model. In IR imaging there is no sign of bulge, we then interpret the mass in ExpDisc component as the baryonic mass from lens galaxy.The model requires that the ExpDisc component makes up about 10 per cent of the total mass within the Einstein radius (1.7~kpc for a lens redshift of $z_l = 1$), which is not unreasonable.  We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models need the full extent of the gravitational arc for constraints. While the gravitational arc in IR band suffers severe dust extinction, the compact lensed images from future high resolution radio imaging will be able to confirm this.

@@

Our analysis for B1555+375 shows that in order to derive a precise quantification of the substructure population and other properties from strongly lensed flux-ratio anomalies, it is crucial to have a careful estimation of non-substructure effects, such as galaxy discs and nearby massive satellites. 
%We find that in B1555+375, a lack of knowledge about the lens galaxy mass distribution, combined with the limited information provided by the position of compact lensed images, can lead to improper lens models and a misinterpretation of the origin of flux-ratio anomalies. 
While upcoming large-scale surveys with the Dark Energy Survey, Kilo-Degree Sky-VIKING, Square Kilometre Array, {\it Euclid}, and Large Synoptic Survey Telescope, are expected to deliver new large samples of gravitationally lensed quasars, in the near future the discovery of new systems using narrow-line quasar emission \citep{N14} will provide a new and invaluable statistical perspective on the current constraints based on flux-ratio anomalies. We stress, therefore, that one should be very careful when interpreting the nature of flux-ratio anomalies, as prior simple assumptions on the lens mass distribution may a significant source of bias. Finally, additional near-infrared AO imaging from SHARP shows that B1555+375 is not the only gravitational lens having a significant disc component detection in the lensing galaxy. 5 out of 25 SHARP lenses are found discs in AO imaging (Fassnacht et al., in prep.).  These gravitational lenses will be investigated in a future work, but the images alone strongly suggest that other systems may have similar links between flux-ratio anomalies and disc components.
